---
title: "ANA 515 Assignment 4 Data Analytics Project"
author: "Prashanth Paisa"
date: "2022-10-15"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Business Goal

The goal of performing analysis of MovieLens Dataset is to create user/movie profile by using their movie watch history. Understanding of the users mindset and according to that recommending them with a movie genre is the important goal of this analysis. Apart from that, this analysis will also help movie industry about the way to release a movie and the genre of the movies from the knowledge extracted from this data analysis.

## Dataset Retrieval

This MovieLens Dataset is retrieved from the following link:

https://drive.google.com/file/d/1Dn1BZD3YxgBQJSIjbfNnmCFlDW2jdQGD/view

## Intializing required libraries

```{r}
library(recommenderlab)
library(ggplot2)
library(data.table)
library(reshape2)
library(Hmisc)
```

## Loading Data

```{r}
movie_data <- read.csv("IMDB/movies.csv",stringsAsFactors=FALSE)
rating_data <- read.csv("IMDB/ratings.csv")
str(movie_data)
```
```{r}
summary(movie_data)
summary(rating_data)
```

## Describing Data

```{r}
describe(movie_data)
describe(rating_data)
```
## Data Preparation

```{r}
movie_genre <- as.data.frame(movie_data$genres, stringsAsFactors=FALSE)
movie_genre2 <- as.data.frame(tstrsplit(movie_genre[,1], '[|]', 
                                   type.convert=TRUE), 
                         stringsAsFactors=FALSE)
colnames(movie_genre2) <- c(1:10)

list_genre <- c("Action", "Adventure", "Animation", "Children", 
                "Comedy", "Crime","Documentary", "Drama", "Fantasy",
                "Film-Noir", "Horror", "Musical", "Mystery","Romance",
                "Sci-Fi", "Thriller", "War", "Western")
genre_mat1 <- matrix(0,10330,18)
genre_mat1[1,] <- list_genre
colnames(genre_mat1) <- list_genre

for (index in 1:nrow(movie_genre2)) {
  for (col in 1:ncol(movie_genre2)) {
    gen_col = which(genre_mat1[1,] == movie_genre2[index,col])
    genre_mat1[index+1,gen_col] <- 1
}
}
genre_mat2 <- as.data.frame(genre_mat1[-1,], stringsAsFactors=FALSE) #remove first row, which was the genre list
for (col in 1:ncol(genre_mat2)) {
  genre_mat2[,col] <- as.integer(genre_mat2[,col]) #convert from characters to integers
} 
str(genre_mat2)
```
## Model Implementation

```{r}
SearchMatrix <- cbind(movie_data[,1:2], genre_mat2[])
ratingMatrix <- dcast(rating_data, userId~movieId, value.var = "rating", na.rm=FALSE)
ratingMatrix <- as.matrix(ratingMatrix[,-1]) #remove userIds
#Convert rating matrix into a recommenderlab sparse matrix
ratingMatrix <- as(ratingMatrix, "realRatingMatrix")
recommendation_model <- recommenderRegistry$get_entries(dataType = "realRatingMatrix")
lapply(recommendation_model, "[[", "description")
recommendation_model$IBCF_realRatingMatrix$parameters

```

## Data Visualization

Here a bar plot has been developed with Trips by Hour and Month. 

```{r}
similarity_mat <- similarity(ratingMatrix[1:4, ],
                               method = "cosine",
                               which = "users")
movie_views <- colCounts(ratingMatrix) # count views for each movie
table_views <- data.frame(movie = names(movie_views),
                          views = movie_views) # create dataframe of views
table_views <- table_views[order(table_views$views,
                                 decreasing = TRUE), ] # sort by number of views
table_views$title <- NA
for (index in 1:10325){
  table_views[index,3] <- as.character(subset(movie_data,
                                         movie_data$movieId == table_views[index,1])$title)
}
table_views[1:6,]
ggplot(table_views[1:6, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = 'steelblue') +
  geom_text(aes(label=views), vjust=-0.3, size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Total Views of the Top Films")
```
